---
title: "Introduction to rsmRobust"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rsmRobust}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 5
)
```

# Overview

**rsmRobust** provides a complete workflow for *robust multiobjective optimization* using:

- Response Surface Methodology (RSM)
- Derringer–Suich desirability
- Sequential first- and second-order models
- Bootstrap confidence intervals for stationary points
- Robustness plots for uncertainty visualization

Based on:

Gamero-Salinas, J., & López-Fidalgo, J. (2025).  
*Response Surface Methodology using desirability functions for multiobjective optimization to minimize indoor overheating hours and maximize useful daylight illuminance.*  
Scientific Reports 15, 12173.  
https://doi.org/10.1038/s41598-025-96376-x

```{r message=FALSE}
library(rsmRobust)
```

---

# 1. Screening Design

```{r}
screen <- run_fractional_screening(
  nfactors = 4,
  resolution = 4,
  factor_names = c("A","B","C","D")
)
screen
```

---

# 2. Noise / Nuisance Factors

```{r}
set.seed(123)
noise <- add_noise_factors(
  n = 10,
  mean = 0,
  sd = 2,
  transform_fn = function(x) ((x * 2) + 15)/100
)
noise
```

---

# 3. Multiresponse Desirability

```{r}
df_demo <- data.frame(
  IOH = c(0.05, 0.15, 0.18),
  UDI = c(80, 50, 40)
)

compute_desirability(
  df_demo,
  ioh_range = c(0,0.2),
  udi_range = c(35,100)
)
```

---

# 4. Sequential RSM

## 4.1 Build CCD and simulate data

```{r}
dsg <- ccd(2, n0 = c(2,2), alpha = "rotatable", randomize = FALSE)

set.seed(1)
true_fun <- function(x1, x2){
  5 + 0.5*x1 - 0.3*x2 - 0.4*x1^2 - 0.6*x2^2 + 0.2*x1*x2
}

dsg$y <- with(dsg, true_fun(x1,x2) + rnorm(nrow(dsg), sd=0.1))
```

## 4.2 Fit second-order model

```{r}
res <- sequential_rsm(
  data = dsg,
  formula = y ~ SO(x1, x2),
  return_real = FALSE
)

res$stationary_coded
res$prediction
```

---

# 5. Bootstrap Confidence Intervals

```{r}
# Build CCD (central composite design)
dsg <- ccd(2, n0 = c(2,2), alpha = "rotatable", randomize = FALSE)

# Simulate a quadratic response surface
set.seed(1)
true_fun <- function(x1, x2){
  5 + 0.5*x1 - 0.3*x2 - 0.4*x1^2 - 0.6*x2^2 + 0.2*x1*x2
}

dsg$y <- with(dsg, true_fun(x1, x2) + rnorm(nrow(dsg), sd = 0.1))

# Fit RSM model
model <- rsm(y ~ SO(x1, x2), data = dsg)

# Bootstrap stationary point
boot <- bootstrap_stationary_point(model, B = 50)
```

---

# 6. Robustness Plot

## 6.1 Align names

```{r}
stationary_point <- rsm::xs(model)

colnames(boot$boot) <- names(stationary_point)
names(boot$ci)      <- names(stationary_point)
```

## 6.2 Plot

```{r}
plot_robustness_2d(
  boot_df   = boot$boot,
  stationary = stationary_point,
  xvar = names(stationary_point)[1],
  yvar = names(stationary_point)[2],
  ci   = boot$ci
)
```


